---
title: "Iris explorer"
author: "Lisa"
date: "`r Sys.Date()`"
output:
  bookdown::html_document2:
    df_print: paged
    toc: true
    toc_float: true
    number_sections: true
    code_folding: hide
    theme: cosmo
---

<style>

ul.task, ul.hint {
  list-style: none;
  margin-left: 0;
  padding-left: 0;
}
ul.task > li, ul.hint > li {
  padding-left: 1em;
  text-indent: -1em;
}

ul.hint > li:before {
  content: "\01F4A1";
  padding-right: 0.25em;
}
ul.task > li:before {
  content: "\01F9E0";
  padding-right: 0.25em;
}

</style>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
rm(list = ls())
```

# Preface

- Scope and goals
- Prerequisites
- Procedure


# The Challenges

## Read the data set

**The data set is stored in `raw_data/irisData.csv`.**

<ul class="task">
  <li>Read the data and store it in a (preliminary) data structure that seems appropriate to you.</li>
  <li>Which data structure do you chose and why?</li>
</ul>

**Hint**

<ul class="hint">
  <li>`csv` is a format for text files.</li>
  <li>The first row of the data file contains the column headings.</li>
  <li>Rows are separated by a line break`\n`.</li>
  <li>Columns are separated by a comma `,`.</li>
</ul>

```{r}
# Read data and store it as "flowers" ----

# Set working directory to the parent directory of "raw_data"
# ...

data_path <- file.path("raw_data", "irisData.csv")
if (file.exists(data_path)) {
  # read the data from the folder and store it as "flowers"
  flowers <- read.csv(data_path,
                      header = TRUE)
  # read.csv stores the data as data.frame, see ?data.frame
} else {
  # hint: iris is a built-in data set that comes with your R installation
  data("iris") # load
  flowers <- iris # name it "flowers"
}

```

## The iris flower

**Do an online research.**

<ul class="task">
  <li>What is an iris? How does it look like? Where do they grow?</li>
  <li>Find pictures of the species _iris setosa_, _versicolor_, and _virginica_.</li>
  <li>Compare the species of irises: What do they have in common, which features distinguish them?</li>
</ul>

**Hint**

<ul class="hint">
  <li>It's fine to use a search engine.</li>
  <li>It's fine to study Wikipedia articles.</li>
</ul>


## Inspect your data, 1

**Get an overview over the data.**

<ul class="task">
  <li>How many rows are there, how many columns?</li>
  <li>What could the columns mean? What could the rows indicate?</li>
  <li>Which kind of values do you find in each column? Which data type would you use to represent them adequately?</li>
</ul>


## Prepare the data

**Data types**

<ul class="task">
  <li>If necessary, convert the data columns to the adequate data type.</li>
  <li>How could you summarize and visualize the data per column? Collect ideas.</li>
</ul>


## Scales of measurement

<ul class="task">
  <li>Find out about the different scales of measurement.</li>
  <li>Which scale of measurement do you think describes the data in each column the best?</li>
  <li>Do your currently used data types reflect the measurement scale? Explain why or why not and convert the data if necessary.</li>
</ul>


## Inspect your data, 2

<ul class="task">
  <li>Make a summary for each column. Which kind of measure or plot type summarizes the data adequately? Explain your choices.</li>
</ul>


## TBC

More ideas&hellip;


# Wrap-up

- Which data analysis skills did you practice?
- Which things did you learn?
- What was easy for you, what was difficult?
- What else could one explore with this data set?
- What would you like to learn next?